---
title: "Bootstrap_Functions"
author: "R. Travis Merrill"
date: "2022-12-24"
output: github_document
---

```{r setup, include=FALSE, echo=FALSE}
library(dplyr)
library(gt)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
opts_chunk$set(echo = TRUE)
```

## Distribution Functions 

```{r, echo=FALSE}
#Inverse cumulative distribution function.
#Returns the value of a desired percentile from a given dataset. 
icdf = function(data,a){
  if(a <= 0.5){
    i = floor(a*(length(data)+1))
  }else{
    i = (length(data)+1) - floor((1-a)*(length(data)+1))
  }
  data = sort(data)
  return(data[i])
}
```

```{r, echo=FALSE}
#Cumulative Distribution Function.
#This reports the proportion of values that are less than or equal to a given value.
cdf = function(data,a){
  data = sort(data)
  vec = which(data <= a)
  b = (length(vec)/length(data))
  return(b)
}
```

##Estimating the population distribution when considering population size. 
<br>

#Large population: Generally N/n > 20, we sample *with* replacement from a *single* copy of the sample
```{r, echo=FALSE}

#data - a data frame
#n - the sample size 
#b- the number of bootstrap iterations, typically 1000+
#This particular function draws a series of *b* random samples of size *n* from the original sample data and
#Determines the median of each RS. This develops the bootstrap distribution of the population median.
#This function reports the mean of this distribution, which provides a statistically sound estimate of the unknown population median. 
#We also report the standard error of the distribution, the bias of the estimator, and Normal, Symetrric, and Non-Symetric confidence intervals of the estimate. 

data = data;

bootmedplus = function(data,n,b){
  BTmedian = rep(0,b);
  for (i in 1:b) {
  BTsamp = sample(data, n, replace = T)
  BTmedian[i] = icdf(BTsamp,0.5)
  };
  BTmedian = sort(BTmedian);
  Bbtmedmean = mean(BTmedian);
  BbtSE = sqrt((1/(b-1))*sum((BTmedian - mean(BTmedian))**2))
  #Normal population distribution
  NormalCI = c(icdf(data,0.5)-BbtSE*1.645, icdf(data,0.5) + BbtSE*1.645)
  #Symmetric population distribution 
  SCI = c(BTmedian[0.05*b], BTmedian[0.95*b + 1])
  #Non-symmetric (no assumption) on pop distribution 
  NSCI = c(2*icdf(data,0.5) - BTmedian[0.95*b + 1], 2*icdf(data,0.5) - BTmedian[0.05*b])
  bias = mean(BTmedian) - icdf(data,0.5)
  output = list("SE" = BbtSE,"bias" = bias,"NormalCI" = NormalCI,"SCI" = SCI,"NSCI" = NSCI)
  return(output)
}
```

##Small population with N/n = a < 20 and an *integer*. We sample *without* replacement from *a* copies of the sample. 
```{r, echo=FALSE}

#data = rep(data, a)

#This particular function estimates the IQR from multivariate data sets.
#It holds each observation's combination of covariates as it samples. 
#n and b are interpreted as before.

bootIQR = function(data,n,b){
  bootiqr = rep(0,b)
  for (i in 1:b) {
  samp = data[sample(seq_len(nrow(data)),n,replace = TRUE),]
  bootiqr[i] = icdf(samp[,1],0.75)-icdf(samp[,1],0.25)
  }
  return(bootiqr)
}

```

#For populations such that N/n = b < 20, b *not* an integer, we use the methods of Booth, Butler, and Hall (1994). The idea is to randomly adjust the composition of the population estimate several times in an attempt to balance out case representation, and avoid bias. 

```{r, echo=FALSE}
#BT is the number of blocks, typically 10
#m is the number of samples within each block 
#This particular function determines an estimate of the population distribution of medians, and returns variase statistics and confidence intervals. 

BBHmedplus = function(data,N,n,BT,m){
  b = BT*m;
  BTmed = c();
  a = floor(N/n);
  k = N - (n*a);
  for (j in 1:BT){
    blockpop = c(rep(data,a),sample(data,k, replace = F));
    for (i in 1:m) {
      blocksamp = c(sample(blockpop,n));
      BTmed = c(BTmed,icdf(blocksamp,0.5));
    }
  }
  BTmed = sort(BTmed)
  Bbtmedmean = mean(BTmed)
  BbtSE = sqrt((1/(b-1))*sum((BTmed - mean(BTmed))**2))
  NormalCI = c(icdf(data,0.5)-BbtSE*1.645, icdf(data,0.5) + BbtSE*1.645)
  SCI = c(BTmed[0.05*b], BTmed[0.95*b + 1])
  NSCI = c(2*icdf(data,0.5) - BTmed[0.95*b + 1], 2*icdf(data,0.5) - BTmed[0.05*b])
  bias = mean(BTmed) - icdf(data,0.5)
  output = list("SE" = BbtSE,"bias" = bias,"NormalCI" = NormalCI,"SCI" = SCI,"NSCI" = NSCI)
  return(output)
}

```


### Examples using functions.   

```{r}
#This synthesized examples pertain to a sample of 7 laboratory mice. Suppose these scores pertain to the post surgery survival time, in days, of mice after given pre-surgical treatment.


micedata = c(94, 197, 16, 38, 99, 141, 23);


bootmedplus(micedata,7,2500)
#Suppose we know that N = 28. Then we have 

micedata1 = rep(micedata,28/7)

bootmedplus(micedata1,7,2500)

#Now suppose that N = 31. Then we have 
BBHmedplus(micedata, 31, 7, 10, 250)

#Notice the similarities in the outputs of these two methods. 
```

